Component response rate variation drives stability in large complex systems

### 2019-03-01 notebook/SI.Rmd
Some work on the SI

- [Stability across increasing $S$](#IncrS) 
- [Stability given targeted manipulation of $\gamma$ (genetic algorithm)](#ga)
- [Stability of ecological networks](#ecological) 
    - [Competitor networks](#competition) 
    - [Mutualist networks](#mutualism) 
    - [Predator-prey networks](#pred-prey) 
- [Sensitivity of connectance (C) values](#connectance)
    - [C = 0.3](#connect3)
    - [C = 0.5](#connect5)
    - [C = 0.7](#connect7)
    - [C = 0.9](#connect9)
- [Large networks of $C = 0.05$ across $S$ and $\sigma$](#sigma)  
    - [$\sigma$ = 0.3](#sigma3)
    - [$\sigma$ = 0.4](#sigma4)
    - [$\sigma$ = 0.5](#sigma5)
    - [$\sigma$ = 0.6](#sigma6)
- [Sensitivity of distribution of $\gamma$](#gam_dist)
- [Feasibility of complex systems](#Feasibility)

@@

>>>>

- [Stability across increasing $S$](#IncrS) 
- [Stability given targeted manipulation of $\gamma$ (genetic algorithm)](#ga)
- [Stability of ecological networks](#ecological) 
    - [Competitor networks](#competition) 
    - [Mutualist networks](#mutualism) 
    - [Predator-prey networks](#pred-prey) 
- [Sensitivity of connectance (C) values](#connectance)
    - [C = 0.3](#connect3)
    - [C = 0.5](#connect5)
    - [C = 0.7](#connect7)
    - [C = 0.9](#connect9)
- [Large networks of $C = 0.05$ across $S$ and $\sigma$](#sigma)  
    - [$\sigma$ = 0.3](#sigma3)
    - [$\sigma$ = 0.4](#sigma4)
    - [$\sigma$ = 0.5](#sigma5)
    - [$\sigma$ = 0.6](#sigma6)
- [Sensitivity of distribution of $\gamma$](#gam_dist)
- [Feasibility of complex systems](#Feasibility)
- [Consistency with Gibbs et al. (2018)](#Gibbs)

@@

Consistency with Gibbs et al. (2018) {#Gibbs}
===============================================================================

The question that I address in the main text is distinct from that of Gibbs et al [@Gibbs2017], which focused instead on the effect of a diagonal matrix of biological species densities $\mathbf{X}$ on a community matrix $\mathbf{M}$ given a species interaction matrix $\mathbf{A}$. This is modelled as below,

$$\mathbf{M} = \mathbf{XA}$$.

Mathematically, the above is identical to my model in the main text where the system $\mathbf{M}$ is defined by component interaction strengths $\mathbf{A}$ and individual component response rates $\mathbf{\gamma}$,

$$\mathbf{M} = \mathbf{\gamma A}$$.

While I focused on the probability of observing a stable versus unstable system given variation in $\mathbf{\gamma}$ for finite by high system size ($S$), Gibbs et al. [@Gibbs2017] instead focused on the different question concerning the effect of $\mathbf{\gamma}$ on the stability of $\mathbf{M}$ given $\mathbf{A}$ as $S \to \infty$. It is easy to show that my primary results in [the table above](#IncrS) are consistent with Gibbs et al. [@Gibbs2017] by calculating the probability that system stability changes as $S$ increases. Gibbs et al. [@Gibbs2017] predict that the probability that stability is altered should decrease with increasing $S$.

```{r, echo = FALSE}
d_tab    <- dat[,1:7];
p_chnged <- (dat[,6] + dat[,7]) / (dat[,2] + dat[,3]);
gibbtb   <- cbind(d_tab[,1], p_chnged);
colnames(gibbtb) <- c("S", "Pr_changed");
plot(x = gibbtb[,1], y = gibbtb[,2], type = "l");
```

Note, however, that within this finite range of $S$, the probability that an unstable system will be stabilised by $\gamma$ increases as $S$ increases. In other words, the proportion of unstable systems that are stabilised by $\gamma$ increases with $S$ in simulations.

```{r, echo = FALSE}
Pr_stbled <- (dat[,6]) / (dat[,6] + dat[,7]);
gibbtb2   <- cbind(d_tab[,1], Pr_stbled);
colnames(gibbtb2) <- c("S", "Pr_stbled");
plot(x = gibbtb2[,1], y = gibbtb2[,2], type = "l");
```

```

### 2019-03-01 notebook/SI.tex
Some work on the SI

\setlength{\droptitle}{-2em}
  \title{Component response rate variation drives stability in large complex
systems}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Supplemental Information}
  \author{A. Bradley Duthie}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{Biological and Environmental Sciences, University of Stirling, Stirling,
UK, FK9 4LA
\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}}

@@

  \begin{itemize}
  \tightlist
  \item
    \protect\hyperlink{sigma3}{\(\sigma\) = 0.3}
  \item
    \protect\hyperlink{sigma4}{\(\sigma\) = 0.4}
  \item
    \protect\hyperlink{sigma5}{\(\sigma\) = 0.5}
  \item
    \protect\hyperlink{sigma6}{\(\sigma\) = 0.6}
  \end{itemize}
\item
  \protect\hyperlink{gam_dist}{Sensitivity of distribution of
  \(\gamma\)}
\item
  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
\end{itemize}

@@

@@

>>>>

\setlength{\droptitle}{-2em}

  \title{Component response rate variation drives stability in large complex
systems}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Supplemental Information}
  \author{A. Bradley Duthie}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{Biological and Environmental Sciences, University of Stirling, Stirling,
UK, FK9 4LA
\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}}

@@

  \begin{itemize}
  \tightlist
  \item
    \protect\hyperlink{sigma3}{\(\sigma\) = 0.3}
  \item
    \protect\hyperlink{sigma4}{\(\sigma\) = 0.4}
  \item
    \protect\hyperlink{sigma5}{\(\sigma\) = 0.5}
  \item
    \protect\hyperlink{sigma6}{\(\sigma\) = 0.6}
  \end{itemize}
\item
  \protect\hyperlink{gam_dist}{Sensitivity of distribution of
  \(\gamma\)}
\item
  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
\item
  \protect\hyperlink{Gibbs}{Consistency with Gibbs et al. (2018)}
\end{itemize}

@@

\hypertarget{Gibbs}{\section{Consistency with Gibbs et al.
(2018)}\label{Gibbs}}

The question that I address in the main text is distinct from that of
Gibbs et al\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{4}}, which
focused instead on the effect of a diagonal matrix of biological species
densities \(\mathbf{X}\) on a community matrix \(\mathbf{M}\) given a
species interaction matrix \(\mathbf{A}\). This is modelled as below,

\[\mathbf{M} = \mathbf{XA}\].

Mathematically, the above is identical to my model in the main text
where the system \(\mathbf{M}\) is defined by component interaction
strengths \(\mathbf{A}\) and individual component response rates
\(\mathbf{\gamma}\),

\[\mathbf{M} = \mathbf{\gamma A}\].

While I focused on the probability of observing a stable versus unstable
system given variation in \(\mathbf{\gamma}\) for finite by high system
size (\(S\)), Gibbs et
al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{4}} instead
focused on the different question concerning the effect of
\(\mathbf{\gamma}\) on the stability of \(\mathbf{M}\) given
\(\mathbf{A}\) as \(S \to \infty\). It is easy to show that my primary
results in \protect\hyperlink{IncrS}{the table above} are consistent
with Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{4}}
by calculating the probability that system stability changes as \(S\)
increases. Gibbs et
al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{4}} predict that
the probability that stability is altered should decrease with
increasing \(S\).

\includegraphics{SI_files/figure-latex/unnamed-chunk-26-1.pdf}

Note, however, that within this finite range of \(S\), the probability
that an unstable system will be stabilised by \(\gamma\) increases as
\(S\) increases. In other words, the proportion of unstable systems that
are stabilised by \(\gamma\) increases with \(S\) in simulations.

\includegraphics{SI_files/figure-latex/unnamed-chunk-27-1.pdf}

```

@@

\hypertarget{ref-Gibbs2017}{}
4. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
population abundances on the stability of large random ecosystems.
\emph{arXiv} (2017).
